Test of Adaptable User Interfaces

### 2015-07-08 paper.tex
Testing Web and Mobile screenshots in the paper

% =============================================================================
\section{Introduction}
% =============================================================================
Nowadays, with the rise of mobile devices such as notebooks, smartphones, smartwatches and tablets, there is computer systems running everywhere and in different \textit{contexts of use}, which involves three factors: the \textit{platform}, \textit{environment} and \textit{the user}~\cite{ui-eric-plasticity-vs-responsive}. These factors affect the interaction between the user and the system. This interaction happens through the \textit{user interface} (UI), and users have to deal with different UIs adapted to several contexts. Hence the term \textit{adaptable user interfaces}. All this allows to see that today developing UIs can be as complex as developing the core of the system. There are several questions to be considered, among them: quality. \textit{Software Testing} is one of the most used ways in the industry to assure quality~\cite{test-hdr-lydie}. It can be defined as a process, or a series of processes, designed to ensure computer code does what it was designed to do and, conversely, that it does not anything unintended~\cite{test-art-of-testing}. What is the impact of adaptable user interfaces on software testing? Are there already tools to test these adaptable UIs?

Answering these questions motivates the development of this project. In this work, we focus on the scenario of having different UIs for the same system. These UIs can be adapted to different devices or have different versions for the same device. For instance, we used as case study a smart home energy management system prototype with four equivalent versions: three for web and one for mobile environment. This exposes the problem which we focus: the repetitive work of writing and maintaining test scripts for different versions of the UIs of the same system.

Our contribution is the approach of automatically generate test scripts for each version of the UI of the same system from scripts written in a higher-level language. In other words, testers would be able to write test cases for all the versions using abstract descriptions. A tool would be responsible for interpreting these descriptions and generating test scripts ready to be executed for each version of the UI.

This paper is structured as follows. First, the related work is presented, explaining several concepts about test automation, UI testing and tools. Then, we explain the case study used in this research. Next, we show how test scripts can be generated from abstract descriptions using an existing tool. Finally, the conclusions that summarize our current results and proposed perspectives.

% =============================================================================
\section{Related Work}
% =============================================================================
With the popularity of Continuous Integration (CI), a software engineering practice in which isolated changes are immediately tested and reported on when they are added to a larger code base, the need of test automation has increased. Test automation can be achieved by three ways: the generation of test scripts, their execution and the analysis of the results. The scientific community has developed several approaches and tools for automation of regular and UI testing. Some of them will be presented below.

An example of automation for the generation of test scripts is Tobias~\cite{tobias}, a combinatorial tool that instantiates a large set of test sequences from an abstract description. Combinatorial testing allows to express a large set of test cases in a few lines. Tobias receives the abstract test and generates an executable test code containing several possible combinations of execution, according to the specification of the abstract code given.

In the case of execution automation, several tools have been developed focusing on UI testing. There are tools that drive the UI as users do, triggering actions, navigating the system and checking the UI content. For instance, tools for web~\cite{watir, cucumber} and mobile~\cite{appium, selendroid} environments are responsible for interacting with the UI according to the scripts written by the testers. Besides, there are tools~\cite{selenium, monkeytalk} that are also able to generate test scripts, through the process of record and playback. The user interacts with the system, the tool records the actions, thus generating a test script that can be used thereafter for execution automation.

Another interesting approach, proposed in~\cite{test-image-comparison-korea}, uses the idea of record and playblack with the automation of the analysis of the results through image comparison. In this solution, after the execution of a test script, the image of the UI is recorded. It can be used later to verify if changes in the system affected the UI, comparing automatically the old and new images of the UI.

Another related work, presented in~\cite{ui-equilavance-raquel}, is not strictly about testing, but comparing UIs of the same system that are adapted to different contexts. In this approach, the UIs are abstracted to formal models based on graphs and the abstractions can be automatically compared, verifying if  they have the same capabilities and appearance.

% =============================================================================
\section{Case Study}
The case study used in this work is a prototype for a smart home energy management system. This decision is based on the fact that it is a modern concept, little explored in the domain of this research and whose UIs can be adapted to several contexts. A smart home may be defined as a well-designed structure with sufficient access to assets, communication, controls, data, and information technologies for enhancing the occupants' quality of life through comfort, convenience, reduced costs and increased connectivity~\cite{smart-home-harper}. A smart home energy management system is an energy management system applied to smart homes. It allows users to control and monitor the energy consumption in a residence from different devices.

@@

>>>>

% =============================================================================
\section{Introduction}
% =============================================================================
TODO

% =============================================================================
\section{Related Work}
% =============================================================================
TODO

% =============================================================================
\section{Case Study}
% =============================================================================

\marginpar{
\begin{figure}
  \begin{center}
  \includegraphics[width=\marginparwidth]{images/web0_goal-filter.eps}
  \caption{Une figure dans la marge.}
  \label{fig:marginparsample}
  \end{center}  
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=\marginparwidth]{images/mobile_goal-filter.eps}
  \caption{Une figure dans la marge.}
  \label{fig:marginparsample}
  \end{center}  
\end{figure}
}

The case study used in this work is a prototype for a smart home energy management system. This decision is based on the fact that it is a modern concept, little explored in the domain of this research and whose UIs can be adapted to several contexts. A smart home may be defined as a well-designed structure with sufficient access to assets, communication, controls, data, and information technologies for enhancing the occupants' quality of life through comfort, convenience, reduced costs and increased connectivity~\cite{smart-home-harper}. A smart home energy management system is an energy management system applied to smart homes. It allows users to control and monitor the energy consumption in a residence from different devices.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{images/web0_goal-filter.eps}
  \caption{Ajoutez une légende en dessous de chaque image.}
  \label{fig:sample}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{images/mobile_goal-filter.eps}
  \caption{Ajoutez une légende en dessous de chaque image.}
  \label{fig:sample}
\end{figure}
